---
title: "MPN bulk WGS n30 - SNV signature analysis"
output:
  html_document:
    code_folding: show
    df_print: paged
    toc: yes
    toc_float:
      collapsed: yes
    toc_depth: 3
    theme: lumen
  pdf_document:
    toc: yes
    toc_depth: '3'
  word_document:
    toc: yes
    toc_depth: '3'
params:
  code: TRUE
---

# Aim

This analysis is to explore the SNV mutational signatures in the bulk WGS MPN cohort (n30)

# 0. Setup

```{r, include=FALSE}
knitr::opts_chunk$set(echo = params$code)
knitr::opts_chunk$set(message = params$code)
knitr::opts_chunk$set(warning = params$code)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## a. Set ref genome
```{r}
SPECIES = "Hsapiens"
SPECIES_BUILD="hg19"
```

## b. Load libraries
```{r, message=FALSE,error=FALSE,results=FALSE}
knitr::opts_chunk$set(collapse = TRUE)

library("BSgenome.Hsapiens.UCSC.hg19")

library("tidyverse")
library("ggrepel")
library("parallel")
library("magrittr")
library("patchwork")

source("revised_sigs.R")
source("contribution.plot.R")

theme_set(theme_bw())
```

## c. Custom functions
```{r}
  addTotal <-
    function(df) {
      df %>%
        mutate(across(where(is.logical), as.character)) %>%
        bind_rows(summarise(., across(where(is.numeric), ~ sum(.x, na.rm = T)), 
                            across(where(is.character), ~ "Total")))
    }
```

## d. Create signature catalogue 
Use COSMIC_v3.2_SBS_GRCh37.txt and add SBSblood from Machado et al. 2022 https://doi.org/10.1038/s41586-022-05072-7
```{r}
pcawg <- get_pcawg60()
pcawg.row.names <- row.names(pcawg)
# SBSblood signature from Table S8 Machado et al. 2022
lym <- read.table("lympho_sbs_sigs.txt", head = T, stringsAsFactors = FALSE)
pcawg <- cbind(pcawg, SBSblood = lym[, "SBSblood"])

pcawg <- apply(pcawg, MARGIN = 2, as.numeric)
row.names(pcawg) <- pcawg.row.names
```


## e. Data
### i. Clone_driver.key.n104
Clone-level metadata for 104 clones inferred by dpclust for the 30 MPN patients in the bulk WGS cohort.
```{r}
clone_driver.key.n104 <- read_csv("20260410_clone_driver.key.n104.csv")
```
#### 1) Summary
```{r}
clone_driver.key.n104 %>%
  group_by(Patient, category_label) %>%
  summarise(clone_tot = length(cluster_no), HQ.clone_tot = sum(In.vignette == TRUE, na.rm = T), .groups = "drop") %>%
  arrange(desc(category_label)) %>% addTotal()
```

#### 2) Add Patient x cluster name
```{r}
clone_driver.key.n104 %<>%  mutate(cluster_name = paste0(Patient, "_", cluster_no))
```

#### 3) Check for multiple clones within the same patient
PD4775 & PD4947 have multiple 9pLOH containing clones
```{r}
clone_driver.key.n104 %>% filter(!is.na(Clone_abv)) %>% group_by(Patient) %>%  count(Clone_abv) %>%  filter(n>1)
```


```{r}
clone_driver.key.n104 %>% filter(Patient  %in% c("PD4775", "PD4947"))
```

##### i) Add cluster_no suffix to 9pLOH only clones 
```{r}
clone_driver.key.n104 %>%
  mutate(Clone_abv = case_when(cluster_name %in% c("PD4775_cl3", "PD4775_cl4", "PD4947_cl5", "PD4947_cl6") & Clone_abv == "9pLOH" ~ paste0(Clone_abv, ".", cluster_no), TRUE ~ Clone_abv)) %>%
  filter(Patient  %in% c("PD4775", "PD4947"))
```

```{r}
clone_driver.key.n104 %<>%
  mutate(Clone_abv = case_when(cluster_name %in% c("PD4775_cl3", "PD4775_cl4", "PD4947_cl5", "PD4947_cl6") & Clone_abv == "9pLOH" ~ paste0(Clone_abv, "_", cluster_no), TRUE ~ Clone_abv)) 
```



### ii. 96-context mutation matrix n104
96-context mutation matrix for 104 clones inferred by dpclust for the 30 MPN patients in the bulk WGS cohort.
```{r}
mut_mat <- read.delim("dp.best.cl.n104.mat.csv", sep = ",")
```

##### 1) Check cluster names match with clone_driver.key
```{r}
table(colnames(mut_mat)  %in%  clone_driver.key.n104$cluster_name)
```


# 1. Run HDP
## a. Subset matrix (96 x 64/104)
Subset matrix to clones in vignettes and greater than 50 mutations
```{r}
mat.name = "current.gt50"
mc2 <- mut_mat[,colSums(mut_mat)>50 & colnames(mut_mat)  %in%  (clone_driver.key.n104 %>% filter(In.vignette) %>%  .$cluster_name)]

dim(mc2)
```

## b. Run HDP
### i. Setup
```{r}
multi.types <- gsub("_.*", "", colnames(mc2))

write.table(mc2, file = paste0(mat.name, ".mat.csv"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = ",")
hash <- digest::digest(list(read.table(paste0(mat.name, ".mat.csv")), multi.types))
```

### ii. Run
Run hdp with burnin=10000, spacing=500 and n_per_chain=250.
```{r}
hdpsave=sprintf("hdp_%s.RDS",hash)

if(file.exists(hdpsave)){
  hdpres2=readRDS(hdpsave)
}else{
  hdpres2=runHDP_denovo(as.data.frame((mc2)))
  saveRDS(hdpres2,file=hdpsave)
}
```

### iii. Chain QC
```{r}
    par(mfrow=c(2,2), mar=c(4, 4, 2, 1))
    p1 <- lapply(chains(hdpres2), plot_lik, bty="L")
    p2 <- lapply(chains(hdpres2), plot_numcluster, bty="L")
    p3 <- lapply(chains(hdpres2), plot_data_assigned, bty="L")
```

### iv. Component count - n5
```{r}
plot_comp_size(hdpres2, bty="L", lab=c(3, 5, 7))
```


### v. Process HDP results and assign known signatures
Known signatures are assigned if cosine similarity is greater than 0.95
```{r}
 hdp2 <- extract_details_from_hdpres(
   res = hdpres2,
   priorsigs = pcawg,
   mutcount = mc2, threshold = 0.95
 )
```


## c. Check reconstruction error
Per Sample reconstruction error. Highlight samples with counts>200 that have reconstruction cosine similarity < 0.9:
```{r}
df <- data.frame(counts = colSums(mc2), reconstruction_csm = hdp2$sample_csm, label = colnames(mc2))
ggplot(df, aes(x = counts, y = reconstruction_csm, label = label)) +
  geom_point() +
  geom_text_repel(data = df %>% filter(counts > 200 & reconstruction_csm < 0.9)) +
  geom_point(data = df %>% filter(counts > 200 & reconstruction_csm < 0.9), col = "red") +
  ylim(c(0.7, 1)) + geom_vline(xintercept = 200, lty = 3)
```

### i. PD6577_cl1 - No driver TN-ET
```{r}
clone_driver.key.n104 %>%  filter(cluster_name == "PD6577_cl1")
```

```{r}
plot_compare_profiles(profile1 = hdp2$mutcount[,"PD6577_cl1"], profile2 = hdp2$reconstruction["PD6577_cl1", ])
```


# 2. Results
## a. Extracted signatures
### i. Plot components
```{r, fig.width= 7, fig.height = 8}
plot_96_profile(hdp2$signatures, condensed = T)
```

### ii. Component similarity to COSMIC 
```{r dev.args=list(pointsize=12),fig.height=6,fig.width=6}
csm <- cos_sim_matrix(hdp2$signature, pcawg)
csm2 <- csm[, apply(csm, 2, max) > 0.7]

plot_cosine_heatmap(csm2, cluster_rows = TRUE, cluster_cols = TRUE, plot_values = TRUE)
```

### iii. Rename Novel signatures
```{r, fig.width= 7, fig.height = 8}
signatures.mat <- hdp2$signatures

sig.rename.vec <- c("SBSblood" ="SBSblood", "SBS1" = "SBS1",  "N1" = "SBS-B", "N2" = "SBS-A", "N3" = "SBS-C")

colnames(signatures.mat) <- sig.rename.vec[colnames(signatures.mat) ]

plot_96_profile(signatures.mat[,c("SBSblood", "SBS1",  "SBS-A", "SBS-B", "SBS-C")], condensed = T)
```


## d. Signature contribution 
### i. All clones n=64
```{r}
hdp2.tb <- contibutionPlot.data(hdp = hdp2, clone.key = clone_driver.key.n104)
dim(hdp2.tb)
```


#### 1) Rename and group sigs
Signature attributions where the lower bound = 0 are categorised as unassigned
```{r}
sig.col.vec <- rev(c(unassigned = "black", `SBS-C` = "#5A96BF",  `SBS-B` ="#995763", `SBS-A` ="#50A382", SBS1 = "#99968C",  SBSblood = "#D3927C"))

hdp2.tb.anno.df <-
  hdp2.tb %>%
  mutate(Signature.orig = Signature) %>%
  mutate(Signature = factor(case_when(
    Signature.orig == "N1" & Contribution_lb95 > 0 ~ "SBS-B",
    Signature.orig == "N2" & Contribution_lb95 > 0 ~ "SBS-A",
    Signature.orig == "N3" & Contribution_lb95 > 0 ~ "SBS-C",
    Contribution_lb95 == 0 ~ "unassigned", TRUE ~ Signature.orig
  ), levels = names(sig.col.vec))) %>% 
  mutate(Sig_type = case_when(Signature  %in%  c("SBS1", "SBSblood") ~ "Endogenous", TRUE ~ "Novel"))
```


#### 2) Relative and absolute Barplot
```{r dev.args=list(pointsize=8),fig.height=12,fig.width=12}
n64.con_barplot <-
  contibutionPlot.con_plot(tb = hdp2.tb.anno.df, facet = "category_label") + scale_fill_manual(values = sig.col.vec) + scale_colour_manual(values = sig.col.vec)
n64.abs_barplot <-
  contibutionPlot.abs_plot(tb = hdp2.tb.anno.df, facet = "category_label") + scale_fill_manual(values = sig.col.vec) + scale_colour_manual(values = sig.col.vec)


n64.con_barplot +
  plot_spacer() +
  n64.abs_barplot +
  plot_layout(widths = c(0.4, 0.0005, 0.6), guides = "collect")
```

#### 3) Clone level signature relative contributions
Signature attributions where the lower bound = 0 are not plotted
```{r, fig.height=5,fig.width=4}

  hdp2.tb.anno.df %>%
  filter(Contribution_lb95 > 0) %>%
  ggplot(aes(x = Signature, y = Contribution, color = Signature)) +
  ggbeeswarm::geom_quasirandom(size = 1.5) +
  facet_grid(. ~ Sig_type, scales = "free_x", space = "free_x") +
  scale_colour_manual(values = sig.col.vec) +
  theme_bw() +
  labs(y = "Relative contribution", x = NULL) +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_blank()
  )
```


### ii. Novel signature high burden clones n=6
#### 1) Select high exposure samples (> 0.20) and untreated patient clones (PD9478) n8
```{r}
outlier.exposure.list <-
  list(
    "Aza" = rownames(hdp2$contr)[hdp2$contr[, "N2"] > 0.2],
    "Hydroxycarbamide" = rownames(hdp2$contr)[hdp2$contr[, "N1"] > 0.2],
    "SBS-C" = rownames(hdp2$contr)[hdp2$contr[, "N3"] > 0.2],
    "Untreated" = grep("PD9478", colnames(mc2), value = T)
  )

outlier.exposure.list
```

##### a) Plot spectra
Plot clone level spectra of outlier and untreated n8
```{r fig.height = 8, fig.width=6}
plot_96_profile(mc2[,unlist(outlier.exposure.list)], ymax = 0.1)
```

##### b) Subset outlier/untreated clones 
```{r}
outlier.tb <-
  right_join(
    hdp2.tb.anno.df,
    unlist(lapply(names(outlier.exposure.list), function(x) {
      d <-
        outlier.exposure.list[[x]]
      names(d) <-
        rep(x, length(d))
      d
    })) %>% enframe(name = "Group", value = "Sample")
  )

outlier.tb %<>% mutate(Group = factor(as.character(Group),
  levels = names(outlier.exposure.list)
))

outlier.tb %<>% mutate(Clone_abv = factor(as.character(Clone_abv),
  levels = rev(c(
    "PD4007_CSF3R/ATRX/PHF6",
    "PD7281_No driver", 
    "PD6607_TP53/RUNX1/EZH2/ZRSR2/ETV6", 
    "PD6623_TP53/17p/chromoplexy", 
    "PD6607_RUNX1/13p", 
    "PD4947_ASXL1/9pLOH", 
    "PD9478_DNMT3A", 
    "PD9478_JAK2exon12"
  ))
))

outlier.tb %>%  select(Patient, category_label, Clone_abv) %>% distinct()
```

#### 2) Relative and absolute Barplot
```{r fig.height=4,fig.width=8}
outlier.con_barplot <-
  contibutionPlot.con_plot(tb = outlier.tb, facet = "Group") + scale_fill_manual(values = sig.col.vec) + scale_colour_manual(values = sig.col.vec)
outlier.abs_barplot <-
  contibutionPlot.abs_plot(tb = outlier.tb, facet = "Group") + scale_fill_manual(values = sig.col.vec) + scale_colour_manual(values = sig.col.vec)

outlier.con_barplot +
  plot_spacer() +
  outlier.abs_barplot +
  plot_layout(widths = c(0.4, 0.0005, 0.6), guides = "collect")
```


# 3. SessionInfo
```{r}
sessionInfo()
```




